---
title: "Project 1"
author: "David Weck"
date: "6/6/2020"
output:
     rmarkdown::github_document:
       toc: true
       toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#Loading required packages
library(tidyverse)
library(httr)
library(jsonlite)
```

# About JSON  
  
JSON (short for JavaScript Object Notation) is a format that is used to store and transfer data. It is text-only format that mirrors the syntax of JavaScript objects, which allows it to be conveniently read and written by people. It is comprised of key/value pairs and arrays, which are structures ubiquitous across nearly all programming languages. Additionally, JSON requires a relatively small amount of computer overhead and can be interpreted by virtually all languages. It is these characteristics that make JSON format an excellent way to store and transfer data. [(Source)](https://www.json.org/json-en.html) All of these characteristics also make JSON a very widely used method for data exchange. Many web applications use JSON to transfer data back and forth with servers, and APIs often store data in JSON format to allow for easy retrieval from users. As mentioned earlier, nearly all programming languages support JSON, further allowing widespread use of the format. [(Source)](https://www.tutorialspoint.com/json/json_overview.htm)
  
# R Packages for JSON  
  
There are 3 main packages for reading JSON data into R: `rjson`, `RJSONIO`, and `jsonlite`. Information on `rjson` and `RJSONIO`, respectively, can be found [here](https://cran.r-project.org/web/packages/rjson/rjson.pdf) and [here](https://cran.r-project.org/web/packages/RJSONIO/RJSONIO.pdf). In this section, I will be focusing on `jsonlite`. `jsonlite` is an R package with efficient tools for reading and creating JSON data. I have chosen it because it is one of the best packages available for communicating with and retrieving data from APIs through R. One of the most important functions in `jsonlite` is `fromJSON`. This function allows for smooth and fast retrieval of JSON data from a JSON string, URL, or file. Another important function, which is also an argument to `fromJSON`, is `flatten`. In JSON data, it is common to see data frames within other data frames. The `flatten` function/argument converts JSON data with this structure into the 2D structure usually desired for data frames. There are many other functions that make `jsonlite` an excellent package for reading JSON data into R. More information about `jsonlite` can be found [here](https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf)
  
# Connecting to NHL API  
  
This section creates functions to pull data from the NHL API and uses those functions to create data frames.  
  
```{r franchise}

#Function to call /franchise
get_franchise <- function(){
  
  #Initializing variables and creating API call
  base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
  end <- 'franchise' #Setting end of url
  call <- paste0(base_url, end) #Pasting together
  
  #Creating get request, extracting text, and converting to R object
  get_req <- GET(call)
  text <- content(get_req, as = 'text')
  json <- fromJSON(text, flatten = TRUE)
  
  #Returning final data frame
  return(as.data.frame(json))
}

```
  
``` {r franchise totals}

#Function to call /franchise-team-totals
get_franchise_totals <- function(){
  
  #Initializing variables and creating API call
  base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
  end <- 'franchise-team-totals' #Setting end of url
  call <- paste0(base_url, end) #Pasting together
  
  #Creating get request, extracting text, and converting to R object
  get_req <- GET(call)
  text <- content(get_req, as = 'text')
  json <- fromJSON(text, flatten = TRUE)
  
  #Returning final data frame
  return(as.data.frame(json))
}

```
  
``` {r seasons records}

#Function to call /franchise-season-records
get_season_records <- function(ID){
  
  if(is.numeric(ID) & ID %in% 1:38){
  
    #Initializing variables and creating API call
    base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
    end <- 'franchise-season-records?cayenneExp=franchiseId=' #Setting end of url
    call <- paste0(base_url, end, ID) #Pasting together
  
    #Creating get request, extracting text, and converting to R object
    get_req <- GET(call)
    text <- content(get_req, as = 'text')
    json <- fromJSON(text, flatten = TRUE)
  
    #Returning final data frame
    return(as.data.frame(json))
    
  } else stop('Please enter a valid franchise ID')
}

``` 
  
``` {r goalie records}

#Function to call /franchise-goalie-records
get_goalie_records <- function(ID){
  
  if(is.numeric(ID) & ID %in% 1:38){
  
    #Initializing variables and creating API call
    base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
    end <- 'franchise-goalie-records?cayenneExp=franchiseId=' #Setting end of url
    call <- paste0(base_url, end, ID) #Pasting together
  
    #Creating get request, extracting text, and converting to R object
    get_req <- GET(call)
    text <- content(get_req, as = 'text')
    json <- fromJSON(text, flatten = TRUE)
  
    #Returning final data frame
    return(as.data.frame(json))
    
  } else stop('Please enter a valid franchise ID')
}

```
  
``` {r skater records}

#Function to call /franchise-skater-records
get_skater_records <- function(ID){
  
  if(is.numeric(ID) & ID %in% 1:38){
  
    #Initializing variables and creating API call
    base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
    end <- 'franchise-skater-records?cayenneExp=franchiseId=' #Setting end of url
    call <- paste0(base_url, end, ID) #Pasting together
  
    #Creating get request, extracting text, and converting to R object
    get_req <- GET(call)
    text <- content(get_req, as = 'text')
    json <- fromJSON(text, flatten = TRUE)
  
    #Returning final data frame
    return(as.data.frame(json))
    
  } else stop('Please enter a valid franchise ID')
}

```

``` {r data frames}

#Creating Data Frames from API calls
franchise <- get_franchise()
franchise_totals <- get_franchise_totals()

#Initializing data frames
season_records <- goalie_records <- skater_records <- NULL

#Looping through all team IDs and combining datasets
for(ID in 1:38){
  
  temp1 <- get_season_records(ID)
  season_records <- rbind(season_records, temp1)
  
  temp2 <- get_goalie_records(ID)
  goalie_records <- rbind(goalie_records, temp2)
  
  temp3 <- get_skater_records(ID)
  skater_records <- rbind(skater_records, temp3)
  
}
```
  
# EDA on NHL Data  
  
```{r numerical summary 1}

#Creating table of position code and most goals scored in one game
knitr::kable(table(skater_records$data.positionCode, skater_records$data.mostGoalsOneGame),
             caption = 'Table of Most Goals Scored in One Game by Each Position')

```
  
This table shows that most defenders only ever score 1 goal in a game. We can also see that centers are more likely than other positions to have games where they score a lot of goals.  
  
```{r numerical summary 2}

#Filtering skaters to active skaters only
active <- filter(skater_records, data.activePlayer == TRUE)

#Creating table of position code by team
knitr::kable(table(active$data.franchiseName, active$data.positionCode),
             caption = 'Table of Position Count by Team')
```
  
This table gives a roster breakdown of each team. It shows how many players of each position, excluding goalie, each team has on their current roster.  
  
```{r bar chart}

#Filtering  skater records down to active Lightning forwards with min 300 games
active <- skater_records %>% 
  filter(data.activePlayer == TRUE, 
         data.gamesPlayed >= 300, 
         data.positionCode != 'D',
         data.franchiseId == 31) %>%
  mutate(fullName = paste(data.firstName, data.lastName, sep = ' '), #Merging first and last name
         goalsPerGame = data.goals / data.gamesPlayed)   #Creating goals per game stat

#Creating plot object and bar chart of goals per game
g <- ggplot(active, aes(x = fullName, y = goalsPerGame))
g + geom_bar(aes(fill = fullName), stat = 'identity', show.legend = FALSE) + 
  coord_flip() +
  labs(x = 'Player', y = 'Goals Per Game', 
       title = 'Goals Per Game of Active Lightning Forwards (Min. 300 Games Played)')
```
  
I created goals per game statistic to get an understanding of which players are the most efficient scorers. Steven Stamkos and Nikita Kucherov stand out as the most efficient goal scorers among active Lightning players. Stamkos, remarkably, scores a goal every two games.
  
```{r boxplot}

#Creating box-plot of log penalty minutes by position
#Using log scale because distribution is very skewed
h <- ggplot(skater_records, aes(x = data.positionCode, y = log(data.penaltyMinutes)))
h + geom_boxplot(aes(fill = data.positionCode), show.legend = FALSE) +
  scale_x_discrete(labels = c('Center', 'Defense', 'Left Wing', 'Right Wing')) +
  labs(x = 'Position', y = 'Log of Penalty Minutes', 
       title = 'Boxplots of Log Penalty Minutes by Position')

```
  
This plot displays boxplots of log career penalty minutes by position. I used log penalty minutes because the distribution of penalty minutes is so heavily skewed. As you can see, it appears defenders generally accumulate slightly more penalty minutes in their careers than other positions. Centers appear to accumulate slightly less penalty minutes. The player with the most penalty minutes in his career was a Right Wing.  
  
```{r scatterplot}

#Filtering to include current teams only and creating per game stats
current <- franchise_totals %>%
  filter(is.na(data.lastSeasonId)) %>%
  mutate(penaltyMinutesPerGame = data.penaltyMinutes / data.gamesPlayed,
         winRatio = data.wins / data.gamesPlayed)

#Changing game type into a factor
current$data.gameTypeId <- as.factor(current$data.gameTypeId)
         
#Plotting scatterplot and fitted lines
l <- ggplot(current, aes(x = penaltyMinutesPerGame, y = winRatio))
l + geom_point(aes(color = data.gameTypeId)) +
  geom_smooth(aes(group = data.gameTypeId, color = data.gameTypeId), method = 'lm', se = FALSE) +
  scale_color_discrete(name = 'Game Type', 
                      labels = c('Regular Season', 'Playoff')) +
  labs(x = 'Penalty Minutes per Game', y = 'Franchise Win/Loss Ratio',
       title = 'Franchise Win/Loss Ratio Vs Penalty Minutes per Game')

```
  
From this plot, we can see that in the regular season, franchises with more penalty minutes per game tend to have a lower overall win-loss ratio. The opposite seems tends to be the case in the playoffs, although the correlation is quite low in both cases.
