---
title: "Project 1"
author: "David Weck"
date: "6/6/2020"
output:
     rmarkdown::github_document:
       toc: true
       toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#Loading required packages
library(tidyverse)
library(httr)
library(jsonlite)
```

# About JSON  
  
JSON (short for JavaScript Object Notation) is a format that is used to store and transfer data. It is text-only format that mirrors the syntax of JavaScript objects, which allows it to be conveniently read and written by people. It is comprised of key/value pairs and arrays, which are structures ubiquitous across nearly all programming languages. Additionally, JSON requires a relatively small amount of computer overhead and can be interpreted by virtually all languages. It is these characteristics that make JSON format an excellent way to store and transfer data. [(Source)](https://www.json.org/json-en.html) All of these characteristics also make JSON a very widely used method for data exchange. Many web applications use JSON to transfer data back and forth with servers, and APIs often store data in JSON format to allow for easy retrieval from users. As mentioned earlier, nearly all programming languages support JSON, further allowing widespread use of the format. [(Source)](https://www.tutorialspoint.com/json/json_overview.htm)
  
# R Packages for JSON  
  
There are 3 main packages for reading JSON data into R: `rjson`, `RJSONIO`, and `jsonlite`. Information on `rjson` and `RJSONIO`, respectively, can be found [here](https://cran.r-project.org/web/packages/rjson/rjson.pdf) and [here](https://cran.r-project.org/web/packages/RJSONIO/RJSONIO.pdf). In this section, I will be focusing on `jsonlite`. `jsonlite` is an R package with efficient tools for reading and creating JSON data. I have chosen it because it is one of the best packages available for communicating with and retrieving data from APIs through R. One of the most important functions in `jsonlite` is `fromJSON`. This function allows for smooth and fast retrieval of JSON data from a JSON string, URL, or file. Another important function, which is also an argument to `fromJSON`, is `flatten`. In JSON data, it is common to see data frames within other data frames. The `flatten` function/argument converts JSON data with this structure into the 2D structure usually desired for data frames. There are many other functions that make `jsonlite` an excellent package for reading JSON data into R. More information about `jsonlite` can be found [here](https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf)
  
# Connecting to NHL API  
  
This section creates functions to pull data from the NHL API and uses those functions to create data frames.  
  
```{r franchise}

#Function to call /franchise
get_franchise <- function(){
  
  #Initializing variables and creating API call
  base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
  end <- 'franchise' #Setting end of url
  call <- paste0(base_url, end) #Pasting together
  
  #Creating get request, extracting text, and converting to R object
  get_req <- GET(call)
  text <- content(get_req, as = 'text')
  json <- fromJSON(text, flatten = TRUE)
  
  #Returning final data frame
  return(as.data.frame(json))
}

```
  
``` {r franchise totals}

#Function to call /franchise-team-totals
get_franchise_totals <- function(){
  
  #Initializing variables and creating API call
  base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
  end <- 'franchise-team-totals' #Setting end of url
  call <- paste0(base_url, end) #Pasting together
  
  #Creating get request, extracting text, and converting to R object
  get_req <- GET(call)
  text <- content(get_req, as = 'text')
  json <- fromJSON(text, flatten = TRUE)
  
  #Returning final data frame
  return(as.data.frame(json))
}

```
  
``` {r seasons records}

#Function to call /franchise-season-records
get_season_records <- function(ID){
  
  if(is.numeric(ID) & ID %in% 1:38){
  
    #Initializing variables and creating API call
    base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
    end <- 'franchise-season-records?cayenneExp=franchiseId=' #Setting end of url
    call <- paste0(base_url, end, ID) #Pasting together
  
    #Creating get request, extracting text, and converting to R object
    get_req <- GET(call)
    text <- content(get_req, as = 'text')
    json <- fromJSON(text, flatten = TRUE)
  
    #Returning final data frame
    return(as.data.frame(json))
    
  } else stop('Please enter a valid franchise ID')
}

``` 
  
``` {r goalie records}

#Function to call /franchise-goalie-records
get_goalie_records <- function(ID){
  
  if(is.numeric(ID) & ID %in% 1:38){
  
    #Initializing variables and creating API call
    base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
    end <- 'franchise-goalie-records?cayenneExp=franchiseId=' #Setting end of url
    call <- paste0(base_url, end, ID) #Pasting together
  
    #Creating get request, extracting text, and converting to R object
    get_req <- GET(call)
    text <- content(get_req, as = 'text')
    json <- fromJSON(text, flatten = TRUE)
  
    #Returning final data frame
    return(as.data.frame(json))
    
  } else stop('Please enter a valid franchise ID')
}

```
  
``` {r skater records}

#Function to call /franchise-skater-records
get_skater_records <- function(ID){
  
  if(is.numeric(ID) & ID %in% 1:38){
  
    #Initializing variables and creating API call
    base_url <- 'https://records.nhl.com/site/api/' #Setting Base URL
    end <- 'franchise-skater-records?cayenneExp=franchiseId=' #Setting end of url
    call <- paste0(base_url, end, ID) #Pasting together
  
    #Creating get request, extracting text, and converting to R object
    get_req <- GET(call)
    text <- content(get_req, as = 'text')
    json <- fromJSON(text, flatten = TRUE)
  
    #Returning final data frame
    return(as.data.frame(json))
    
  } else stop('Please enter a valid franchise ID')
}

```

``` {r data frames}

#Creating Data Frames from API calls
franchise <- get_franchise()
franchise_totals <- get_franchise_totals()
season_records <- get_season_records(31)
goalie_records <- get_goalie_records(31)
skater_records <- get_skater_records(31)

```
  
# EDA on NHL Data  
